Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-08-24 Paper_draft/main.tex
Add punctuation to equations
First set of comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\bibliographystyle{iopart-num_custom}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{comment}
\usepackage{acronym}
\usepackage{standalone}
\usepackage{academicons}
\usepackage{scalerel}
\usepackage{layouts}
%\usepackage{import}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,angles,quotes,shapes.geometric, arrows,calc}
\usetikzlibrary{svg.path}
\expandafter\let\csname equation*\endcsname\relax 
\expandafter\let\csname endequation*\endcsname\relax 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[capitalize]{cleveref}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\newcommand{\jordan}[1]{\textbf{\textcolor{red}{JORDAN: #1}}}
\newcommand{\siong}[1]{\textbf{\textcolor{blue}{SIONG: #1}}}
\newcommand{\chris}[1]{\textbf{\textcolor{green}{CHRIS: #1}}}
\newcommand{\michael}[1]{\textbf{\textcolor{orange}{MICHAEL: #1}}}
\newcommand{\dcc}{LIGO-PXXXXXXX}
%\input{tag.tex}
% Command for ORCID id's
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

@@

\begin{abstract}
The next generation of Gravitational wave detectors will accelerate the number
of gravitational wave detection's such that we can gain new in site into the
physics behind the sources causing the phenomena. Numerical simulations and
matched filtering are the standard for detecting gravitational waves for know
sources such as binary-black hole mergers. Other sources of gravitational waves
that remain elusive to standard modelling techniques and are expected to be
detectable, however, there must be a way to characterise them in order to build
a model template. Here we construct a unmodeled burst generation scheme using
Generative adversarial networks - a powerful class of machine learning.
\end{abstract}

@@

% Discuss the family of burst waveforms currently used and why - not in detail, just
% an introduction 
%
\ac{GW} burst algorithms \cite{Klimenko_2008, Aso_2008} are tested and tuned using model waveforms that may
or may not have astrophysical significance but have easy to define parameters
and share characteristics of real bursts that is enough to simulate a \ac{GW} passing between detectors. Such waveforms include sine-Gaussians: a Gaussian modulated sine wave that is by it's central frequency and decay parameter. Bandlimited white-noise bursts: white noise that is contained within a certain frequency range and ring-downs which mimic the damped oscillations after a CBC merger.

@@

With the expectation that there will be many more \ac{GW} detections in the future, researches are starting to realise the need for fast and efficient \ac{GW} analysis to compete with the rising number of detections. While still in its infancy, Machine Learning (ML) with GW has already shown great potential for overcoming data bottlenecks. \ac{GW} analysis in areas of detection \cite{Gabbard2017,Gebhard_2019,Krastev_2020}, glitch classification \cite{Bahaadini, George_2018,Razzano_2018} and parameter estimation \cite{gabbard2019bayesian,shen2019deterministic,green2020gravitationalwave} have shown the success of in ML pattern recognition. Applying these algorithms show similar results to matched filtering techniques and have the benefit of being orders of magnitude faster.

@@

In this work we aim to explore the use of machine learning to generate and interpret
unmodelled \ac{GW} burst waveforms. Using the generative machine learning model: Generative Adversarial Networks (GANs), we train on five classes of known burst morphologies in the time domain. Working on the assumption that GANs construct smooth n-dimension vector spaces between it's input and output, we can then explore the space between the five classes to construct new unmodelled waveforms. As all the computationally expensive processes occur during training, the learned model after training is able to produce waveforms much faster than and in fact produce waveforms that are impossible to produce with current techniques. These new varieties of waveforms can then be used to diagnose detection algorithms and gain new insight into sources of \ac{GW} bursts. 
% the point of the paper
%

@@

Since the birth of programmable computers, people have wondered if machines can develop true intelligence beyond formal mathematical procedures. Today, ML aims to learn apparent relationships held within the given data or `training data' in order to make accurate predictions without the need for additional programming . A common approach in ML relies on the computer learning on past experience to make decisions on future events. The hierarchical nature of this approach means that the machine can build complicated concepts from simpler ones. This branch of AI is called deep learning.
Neural networks are the quintessential machine learning algorithm that aims to approximate a function. They are built from many single processing units called neurons. The simpliest Neural Network is the perceptron layer \cref{fig:perceptron} which holds a single neuron that takes several real inputs $x_{1}, ... , x_{i}$ and maps them to an output according to the following linear equation:

\begin{equation}
\sigma(\sum_i w_i x_i + b)
\label{eqn:neuron}
\end{equation}
where $w$ and $b$ are the weights and bias and $\sigma$ denotes the activation function. The weights are numbers which can be thought as the strength between connections. The output of a neuron is called its activation. Different choices of activation functions allow the user to simulate various models. A common example of an activation function is the hyperbolic tangent \cref{fig:tanh_activation} that is used when the outputs of the network must be both positive and negative and re-scaled to $[-1,1]$. It is often useful to introduce a bias, $b$, such that the neuron remains inactive above zero but is active when the sum reaches a defined threshold. This bias is added before the activation function and it tells us how high the weighted sum needs to be before the neuron becomes active.

@@

A neural network contains many single neurons connected in a layered structure as shown in \cref{fig:neural_network}. The activations of the first layer (or input layer) act as the inputs to the second layer and so on until the output layer. Multilayered neural networks have intermediate layers between the input and output stages dubbed the hidden layers as the computations performed are not accessible to the user. The network can be thought of as a function F: $\mathbb{R}^N\rightarrow \mathbb{R}^M $ reliant on how activations from one layer bring activations in the next. The system is analogous to biology, where, some groups of neurons in animal brains cause certain others to fire.

\subsection{Convolutional Neural Networks}
Convolutional Neural Networks (CNNs) are designed to work with grid-like structures that exhibit strong local spatial dependencies. An example would be a two-dimensional image as each coloured pixel has similar intensities to their near neighbours. The value contained in each pixel can be thought of as adding an additional dimension or channel to the image. Although most work with CNNs involve image-based data, they can be applied to other spatially adjacent data types such as time-series and text items. CNNs are defined by the use of a convolution operation, a mathematical operation that expresses the amount overlap between functions. The convolution is applied by shifting one grid-like structure over the other, drawing out spatially important features between the two.  A strided convolution reduces the dimensionality of the input while retaining spatial features. This can result in a loss of information along the boarders. To combat this, the input is padded with zeros at the edges which does not interfere with the convolutional operation. The output of the convolutional layer is then passed to an activation function.

@@

% How a GAN works
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
that is set up to distinguish between real and fake data and a generator that
produces synthetic reproductions of the real data. The generator performs a
mapping from an input noise vector $\mathbf{z}$, that is usually sampled from a n-dimensional Gaussian space known as the latent space, to its representation of the
data and the discriminator maps its input $\mathbf{x}$ to a probability that
the input came form either the training data or generator.  During training,
the discriminator is given a batch of samples that contains one half real data
and one half fake data which it then makes predictions on. The loss for the
discriminator is calculated by comparing its predictions to the labelled data
through the binary cross-entropy function. \jordan{note to myself: This is important since eqn 2 is derived from it (thats where the logs come from)} The training process of a \ac{GAN}
alternatively updates the weights of the discriminator and generator based on
information on its competitors loss function. This loss of discriminator is
used to update the weights of the generator to produce more realistic samples
of the input distribution, the loss of the generator encourages the
discriminator to update its classification abilities. GANs seek to find an equilibrium between the generator and discriminator in the form of a two-player game that can be summarised by the following:

\begin{comment}
Both networks compete in
a minimax game~\chris{what is a minimax game? references?} which the generator
is trying to minimise and the discriminator is trying to
maximise:~\chris{nicely written, however, it needs an extra 25\% explanation
for people not familar with ML or GANs in particular. Try to spell things out
more. Also refer to Fig 1 somewhere in this section.}
\end{comment}
%
\begin{equation}
\mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x} \sim p_{\text{r}}(\mathbf{x})} [\text{log} D(\mathbf{x})] \\ + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z})))]
\label{equation:GANloss}
\end{equation}

where, V is an objective function that is to be optimized. Objective functions are a general term for either a loss function (to be minimized) or a negative loss function (to be maximised). D is the discriminator, G is the generator and ${\mathbf{x} \sim p_{\text{r}}(\mathbf{x})}$, ${\mathbf{z} \sim p_{\text{z}}(\mathbf{z})}$ are samples taken from the training data and generated data respectfully. In practice this approach was found to be unstable. Early on in training the generators weights are initialised randomly meaning that the early generations are noisy and uninformative. This means that the discriminator can easily learn to spot the fake generations and so the discriminator wins and the generator cannot continue to improve. To overcome this the framing of the generator is often changed. Rather than minimizing $\text{log}(1-D(G(\mathbf{z})))$ the objective now is to maximise $\text{log}(D(G(\mathbf{z})))$, that is, to maximise the probability of the generations being predicted as real. The idea being that we always want the loosing side to be able to leverage from its competitor. 
%

@@

% introduce CGANs and ACGANs
%
To gain more control over what a GAN is able to generate, a conditional variant of \acp{GAN} named \acp{CGAN}~\cite{cgan} was introduced by feeding in extra information into the generator and discriminator such as a class label or attribute label, c. The objective function is then modified:
\begin{equation}
\mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x,c} \sim p_{\text{r}}(\mathbf{x,c})} [\text{log} D(\mathbf{x,c})] \\ + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z}),\mathbf{c} \sim p_{\text{r}}(\mathbf{c})} [\text{log}(1-D(G(\mathbf{z,c})))]
\label{equation:cGANloss}
\end{equation}

@@

GW burst signals remain an unmodelled phenomenon, as such, current
detection algorithms focus on waveforms signals appearing within multiple detectors We propose a
signal generation scheme using \acp{GAN} trained on burst-like
waveforms. BurstGAN is a conditional GAN that is trained on five signal morphology's spanning a range of prior
parameters. The parameter ranges can be seen in \cref{Tab:training_parms} The families are: sine-Gaussian $h(t) = A \exp\left[ - (t-t_{0})^2 / \tau^2 \right] \sin (2 \pi f_0 (t-t_0))$, a sinusoidal wave with a Gaussian envelop characterised by a central frequency, $f_0$; Ring-down, $h(t) = A \exp \left[-{(t-t_0)} / {\tau} \right] \sin(2 \pi f_0 (t-t_0))$, with frequency $f_0$ and duration $\tau$, white-noise bursts, with bounded frequency bandwidth $\Delta f$ and duration $\tau$, gaussian pulse $h(t) = \exp(-t^2 / \tau^2)$ with duration $\tau$ and binary-black hole inspirals which are simulated using IMRPhenomD waveform \cite{Khan_2016} routine from
LALSuite \cite{lalsuite} which models the
inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses lie
in the range of [5,70] $\textrm{M}_{\odot}$ with zero spins
and we fix m$_1$ $>$ m$_2$. The mass distribution is approximated by a power
law with index of 1.6 \cite{Abbott_2019}. The signals are generated using random right ascensions and
declinations uniform over the sky and the inclinations are drawn from the
cosine of a uniform distribution in the range [-1,1]. The peaks of the
waveforms are set to be within [0.4,0.6]s of the 1s time interval.

@@

% Introduce some architecture features
%
Extensions to the original \ac{GAN} method such as the
\ac{DCGAN}~\cite{Radford2015} have been widely praised for
enabling a stable \ac{GAN} architecture. Most GANs now replace fully connected layers, where each neuron is connect to all the neurons in the previous and next layer, with convolutional layers. \ac{CNN} are designed to work with grid-like structures with close local dependencies like image and text based data, however, there are examples of audio synthesis work \cite{DBLP:journals/corr/abs-1809-11096}.

% specifically define our architecture
%
We adopt the suggestions of \cite{Radford2015,DBLP:journals/corr/abs-1809-11096}, lengthening
one-dimensional convolution kernels on both the generator and discriminator.
The Generator model is fully convolutional, upsampled using strided transposed
convolutions with batch normalisation in the first layer and ReLU activations
throughout with the exception of Tanh for the output layer. Each transposed
convolutional layer uses a kernel size of $18\times 1$ and stride of 2. The
discriminator network mirrors that of the generator without batch
normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride
convolution for downsampling. The discriminator output is a single node activated by a Sigmoid that can be interpreted as
a probability of the the signal being real. This model is trained
with binary cross entropy. The full architecture description can be seen in \cref{Tab:hyperparameters}. ~\jordan{still too technical im not sure if we need to go into more detail on each thing mentioned.}
%~\chris{OK, very technically tight but too
%computer sciencey. Try to make it human readbale for a physicist and refer to
%the table of parameters in the appendix.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}

@@

Given a 100 dimensional vector drawn from a normal distribution, a class label and sky localisation information, the GAN is able to generate burst-like waveforms  generalised from the training set. We set out by describing the quality of generated waveforms and how they compare to the training set. We then explore the structure of the latent and class spaces by interpolating between points in these spaces. We test vector arithmetic that can be used to generate a new breed of signal by merging two or more families together. Finally, we discuss the capacity of the discriminator as a \ac{GW} burst classifier and the auxiliary component of this work.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/generations/z_interp_sg.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_rd.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_wnb.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_blip.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_bbh.png}
    \caption{Latent space linear interpolation, class space is held constant.}
    \label{fig:z_interp}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\bibliographystyle{iopart-num_custom}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{comment}
\usepackage{acronym}
\usepackage{standalone}
\usepackage{academicons}
\usepackage{scalerel}
\usepackage{layouts}
%\usepackage{import}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,angles,quotes,shapes.geometric, arrows,calc}
\usetikzlibrary{svg.path}
\expandafter\let\csname equation*\endcsname\relax 
\expandafter\let\csname endequation*\endcsname\relax 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[capitalize]{cleveref}

\usepackage[mathlines]{lineno}% Enable numbering of text and display math
\linenumbers\relax % Commence numbering lines

\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\newcommand{\jordan}[1]{\textbf{\textcolor{red}{JORDAN: #1}}}
\newcommand{\siong}[1]{\textbf{\textcolor{blue}{SIONG: #1}}}
\newcommand{\chris}[1]{\textbf{\textcolor{green}{CHRIS: #1}}}
\newcommand{\michael}[1]{\textbf{\textcolor{orange}{MICHAEL: #1}}}
\newcommand{\dcc}{LIGO-PXXXXXXX}
%\input{tag.tex}
% Command for ORCID id's
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

@@

\begin{abstract}
The next generation of Gravitational wave detectors will accelerate the number
of gravitational wave detections such that we can gain new in site into the
physics behind the sources causing the phenomena. Numerical simulations and
matched filtering are the standard for detecting gravitational waves for known
sources such as binary-black hole mergers. Other sources of gravitational waves
that remain elusive to standard modelling techniques and are expected to be
detectable, however, there must be a way to characterise them in order to build
a model template. \michael{Don't quite follow this sentence.} Here we construct a unmodelled burst generation scheme using
Generative adversarial networks - a powerful class of machine learning.
\end{abstract}

@@

% Discuss the family of burst waveforms currently used and why - not in detail, just
% an introduction 
%
\ac{GW} burst detection algorithms \cite{Klimenko_2008, Aso_2008} are tested and tuned using modelled waveforms that may
or may not have astrophysical significance but have easy to define parameters
and share characteristics of real bursts that are enough to simulate a \ac{GW} passing between detectors. Such waveforms include sine-Gaussians: a Gaussian modulated sine wave that is characterised by its central frequency and decay parameter. Bandlimited white-noise bursts: white noise that is contained within a certain frequency range and ring-downs which mimic the damped oscillations after a CBC merger.

@@

With the expectation that there will be many more \ac{GW} detections in the future, researchers are starting to realise the need for fast and efficient \ac{GW} analysis to compete with the rising number of detections. While still in its infancy, Machine Learning (ML) with GW has already shown great potential for overcoming data bottlenecks. \ac{GW} analysis in areas of detection \cite{Gabbard2017,Gebhard_2019,Krastev_2020}, glitch classification \cite{Bahaadini, George_2018,Razzano_2018} and
parameter estimation \cite{gabbard2019bayesian,shen2019deterministic,green2020gravitationalwave} have shown the success of in ML pattern recognition. Applying these algorithms show similar results to matched filtering techniques and have the benefit of being orders of magnitude faster. \michael{We mention aspects related to detection (pattern matching) but also metion PE in the same sentence and I think this is a bit confusing.}

@@

In this work we aim to explore the use of machine learning to generate and interpret
unmodelled \ac{GW} burst waveforms. Using the generative machine learning model: Generative Adversarial Networks (GANs), we train on five classes of known burst morphologies in the time domain. Working with the assumption that GANs construct smooth n-dimensional vector spaces between its input and output, we can then explore the space between the five classes to construct new unmodelled waveforms. As all the computationally expensive processes occur during training, the learned model after
training is able to produce waveforms much faster than \michael{than what?} and in fact produce waveforms that are impossible to produce with current techniques. These new varieties of waveforms can then be used to diagnose detection algorithms and gain new insight into sources of \ac{GW} bursts. 
% the point of the paper
%

@@

Since the birth of programmable computers, people have wondered if machines can develop true intelligence beyond formal mathematical procedures. Today, ML aims to learn apparent relationships held within the given data or `training data' in order to make accurate predictions without the need for additional programming . A common approach in ML relies on the computer learning on past experience to make decisions on future events. \michael{Not sure about the use of computer here, maybe
\textit{relies on a model}?} The hierarchical nature of this approach means that the machine can build complicated concepts from simpler ones. This branch of AI is called deep learning.
Neural networks are the quintessential machine learning algorithm that aims to approximate a function. They are built from many single processing units called neurons. The simpliest Neural Network is the perceptron layer \cref{fig:perceptron} which holds a single neuron that takes several real inputs $x_{1}, ... , x_{i}$ and maps them to an output according to the following linear equation:

\begin{equation}
\sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{equation}
where $w$ and $b$ are the weights and bias and $\sigma$ denotes the activation function. The weights are numbers which can be thought as the strength between connections. The output of a neuron is called its activation. Different choices of activation functions allow the user to simulate various models. A common example of an activation function is the hyperbolic tangent \cref{fig:tanh_activation} that is used when the outputs of the network must be both positive and negative and re-scaled to $[-1,1]$. It is often useful to introduce a bias, $b$, such that the neuron remains inactive above zero but is active when the sum reaches a defined threshold. This bias is added before the activation function and it tells us how high the weighted sum needs to be before the neuron becomes active.

@@

A neural network contains many single neurons connected in a layered structure as shown in \cref{fig:neural_network}. The activations of the first layer (or input layer) act as the inputs to the second layer and so on until the output layer. Multi-layered neural networks have intermediate layers between the input and output stages dubbed the hidden layers as the computations performed are not accessible to the user. The network can be thought of as a function F: $\mathbb{R}^N\rightarrow \mathbb{R}^M $ reliant on how activations from one layer bring activations in the next. The system is analogous to biology, where, some groups of neurons in animal brains cause certain others to fire.

\subsection{Convolutional Neural Networks}
Convolutional Neural Networks (CNNs) are designed to work with grid-like structures that exhibit strong local spatial dependencies. An example would be a two-dimensional image as each coloured pixel has similar intensities to their near neighbours. The value contained in each pixel can be thought of as adding an additional dimension or channel to the image. Although most work with CNNs involve image-based data, they can be applied to other spatially adjacent data types such as time-series and
text items. CNNs are defined by the use of a convolution operation, a mathematical operation that expresses the amount overlap between functions. The convolution is applied by shifting one grid-like structure over the other, drawing out spatially important features between the two.  A strided convolution reduces the dimensionality of the input while retaining spatial features \michael{How? This is core part of the architecture of the GAN but is quite vague, I don't think it's
obvious that stride imples moving the filter further. Maybe something like: \textit{the distance by which the grid is shifted is known as the stride and increasing it reduces the dimensionality of the output}}. This can result in a loss of information along the boarders. To combat this, the input is padded with zeros at the edges which does not interfere with the convolutional operation. The output of the convolutional layer is then passed to an activation function.

@@

% How a GAN works
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
that is set up to distinguish between real and fake data and a generator that
produces synthetic reproductions of the real data. \michael{Is it clear to the reader that the \textit{fake} data is the same as the synthetic data?} The generator performs a
mapping from an input noise vector $\mathbf{z}$, that is usually sampled from a n-dimensional Gaussian space known as the latent space, to its representation of the
data and the discriminator maps its input $\mathbf{x}$ to a probability that
the input came form either the training data or generator.  During training,
the discriminator is given a batch of samples that contains one half real data
and one half fake data which it then makes predictions on. The loss for the
discriminator is calculated by comparing its predictions to the labelled data
through the binary cross-entropy function. \jordan{note to myself: This is important since eqn 2 is derived from it (thats where the logs come from)} The training process of a \ac{GAN}
alternatively updates the weights of the discriminator and generator based on
information on its competitors loss function. This loss of discriminator is
used to update the weights of the generator to produce more realistic samples
of the input distribution, the loss of the generator encourages the
discriminator to update its classification abilities. GANs seek to find an equilibrium between the generator and discriminator in the form of a two-player game that can be summarised by the following:

\begin{comment}
Both networks compete in
a minimax game~\chris{what is a minimax game? references?} which the generator
is trying to minimise and the discriminator is trying to
maximise:~\chris{nicely written, however, it needs an extra 25\% explanation
for people not familar with ML or GANs in particular. Try to spell things out
more. Also refer to Fig 1 somewhere in this section.}
\end{comment}
%
\begin{equation}
  \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x} \sim p_{\text{r}}(\mathbf{x})} [\text{log} D(\mathbf{x})] \\ + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z})))],
\label{equation:GANloss}
\end{equation}

where V is an objective function that is to be optimized. Objective functions are a general term for either a loss function (to be minimized) or a negative loss function (to be maximised). D is the discriminator, G is the generator and ${\mathbf{x} \sim p_{\text{r}}(\mathbf{x})}$, ${\mathbf{z} \sim p_{\text{z}}(\mathbf{z})}$ are samples taken from the training data and generated data respectfully. In practice this approach was found to be unstable \michael{Cite?}. Early on in training the
generators weights are initialised randomly meaning that the early generations are noisy and uninformative. This means that the discriminator can easily learn to spot the fake generations and so the discriminator wins and the generator cannot continue to improve. To overcome this the framing of the generator is often changed. Rather than minimizing $\text{log}(1-D(G(\mathbf{z})))$ the objective now is to maximise $\text{log}(D(G(\mathbf{z})))$, that is, to maximise the probability of the
generations being predicted as real. The idea being that we always want the loosing side to be able to leverage \michael{information?} from its competitor. 
%

@@

% introduce CGANs and ACGANs
%
To gain more control over what a GAN is able to generate, a conditional variant of \acp{GAN} named \acp{CGAN}~\cite{cgan} was introduced by feeding in extra information into the generator and discriminator such as a class label or attribute label, c. The objective function \cref{equation:GANloss}  is then modified:
\begin{equation}
  \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x,c} \sim p_{\text{r}}(\mathbf{x,c})} [\text{log} D(\mathbf{x,c})] \\ + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z}),\mathbf{c} \sim p_{\text{r}}(\mathbf{c})} [\text{log}(1-D(G(\mathbf{z,c})))].
\label{equation:cGANloss}
\end{equation}

@@

GW burst signals remain an unmodelled phenomenon, as such, current
detection algorithms focus on waveforms signals appearing within multiple detectors We propose a
signal generation scheme using \acp{GAN} trained on burst-like
waveforms. BurstGAN is a conditional GAN that is trained on five signal morphology's spanning a range of prior
parameters. The parameter ranges can be seen in \cref{Tab:training_parms}. The families are: sine-Gaussian $h(t) = A \exp\left[ - (t-t_{0})^2 / \tau^2 \right] \sin (2 \pi f_0 (t-t_0))$, a sinusoidal wave with a Gaussian envelop characterised by a central frequency, $f_0$; Ring-down, $h(t) = A \exp \left[-{(t-t_0)} / {\tau} \right] \sin(2 \pi f_0 (t-t_0))$, with frequency $f_0$ and duration $\tau$, white-noise bursts, with bounded frequency bandwidth $\Delta f$ and duration $\tau$, gaussian pulse $h(t) = \exp(-t^2 / \tau^2)$ with duration $\tau$ and binary-black hole inspirals which are simulated using IMRPhenomD waveform \cite{Khan_2016} routine from
LALSuite \cite{lalsuite} which models the
inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses lie
in the range of [5,70] $\textrm{M}_{\odot}$ with zero spins
and we fix m$_1$ $>$ m$_2$. The mass distribution is approximated by a power
law with index of 1.6 \cite{Abbott_2019}. The signals are generated using random right ascensions and
declinations uniform over the sky and the inclinations are drawn from the
cosine of a uniform distribution in the range [-1,1]. The peaks of the
waveforms are set to be within [0.4, 0.6]s of the 1s time interval.

@@

% Introduce some architecture features
%
Extensions to the original \ac{GAN} method such as the
\ac{DCGAN}~\cite{Radford2015} have been widely praised for
enabling a stable \ac{GAN} architecture. Most GANs now replace fully connected layers, where each neuron is connect to all the neurons in the previous and next layer, with convolutional layers. \ac{CNN} \michael{capitalisation, not sure how it works with the accronyms package} are designed to work with grid-like structures with close local dependencies like image and text based data, however, there are examples of audio synthesis work \cite{DBLP:journals/corr/abs-1809-11096}.

% specifically define our architecture
%
We adopt the suggestions of \cite{Radford2015,DBLP:journals/corr/abs-1809-11096}, lengthening
one-dimensional convolution kernels on both the generator and discriminator.
The Generator model is fully convolutional, upsampled using strided transposed
convolutions with batch normalisation in the first layer and ReLU activations
throughout with the exception of Tanh for the output layer. Each transposed
convolutional layer uses a kernel size of $18\times 1$ and stride of 2. The
discriminator network mirrors that of the generator without batch
normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride
convolution for downsampling. The discriminator output is a single node activated by a Sigmoid that can be interpreted as
a probability of the the signal being real. This model is trained
with binary cross entropy. The full architecture description can be seen in \cref{Tab:hyperparameters}. ~\jordan{still too technical im not sure if we need to go into more detail on each thing mentioned.} \michael{This para doesn't seem to technical to me}
%~\chris{OK, very technically tight but too
%computer sciencey. Try to make it human readbale for a physicist and refer to
%the table of parameters in the appendix.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{comment}

@@

Given a 100 dimensional vector drawn from a normal distribution, a class label and sky localisation information, the GAN is able to generate burst-like waveforms  generalised from the training set \michael{How is the sky sampled? We know it's uniform but we should say explicitly}. We set out by describing the quality of generated waveforms and how they compare to the training set. We then explore the structure of the latent and class spaces by interpolating between points in these spaces. We test vector arithmetic that can be used to generate a new breed of signal by merging two or more families together. Finally, we discuss the capacity of the discriminator as a \ac{GW} burst classifier and the auxiliary component of this work.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/generations/z_interp_sg.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_rd.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_wnb.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_blip.png}
    \includegraphics[width=\textwidth]{figures/generations/z_interp_bbh.png}
    \caption{Latent space linear interpolation, class space is constant.}
    \label{fig:z_interp}
\end{figure}
